Over-relaxed Sinkhorn's Algorithm for Regularized Optimal Transport

### 2017-10-18 paper.tex
Notation Theta

\begin{proposition}
	We argue that a good choice for the functions $\theta_k$ may be constructed as follows. Pick a target parameter $\theta_0 \in [1;2)$, and a small $\delta>0$.
	Define the function $\theta_k^*$ as:
	\begin{equation}
	\theta_k^*(\gamma) = \sup \left\{\omega \in [1;2]  \mid \varphi_\omega\left(\min \frac{A_k \gamma}{\mu^k}\right) \ge 0 \right\} .
	\end{equation}
	Then the function
	\begin{equation}
	\label{eq:theta_k}
	\theta_k(\gamma) = \min(\max(1,\theta_k^*(\gamma)-\delta),\theta_0)
	\end{equation}
	is continuous, and verifies condition (\ref{eq:cond_theta_k}).
\end{proposition}
This construction has several advantages.
One is that it allows to choose arbitrarily the parameter $\theta_0$ that will be used asymptotically. A heuristic for this choice is proposed in section \ref{section:local}.
But it is also an easy approach to having an adaptive method, as the approximation of $\theta_k^*$ has a negligible cost (it only require to solve a one dimensional problem that depends on the smallest value of $\frac{A_k \gamma}{\mu^k}$, and the corresponding function to invert is decreasing and concave).
We will show later on that this is often an excellent choice in practice.
\begin{proof}
	Since the partial derivative of $\frac{\partial\varphi_\omega(x)}{\partial\omega}$ is nonzero for any $x<1$, the implicit function theorem proves the continuity of $\theta_k^*$.
	The function $\theta_k^*(\gamma)$ is constructed so that every term in relation (\ref{eq:kl_diff_scal}) is non-negative.
	Therefore, by lemma \ref{lemma:F_P_theta}, using parameter $\theta_k^*(\gamma)-\delta$ ensures the strong decrease (\ref{eq:cond_theta_k}) of the Lyapunov function.
	It is no problem to constrain it to $[1,\theta_0]$, as using 1 always validates the inequality (\ref{eq:cond_theta_k}), and using $\theta_0$ when it is smaller than $\theta_k^*(\gamma)$ does as well.
\end{proof}

@@

\renewcommand{\algorithmiccomment}[1]{\hfill\bgroup(#1)\egroup}
%\begin{algorithm}
%\caption{Over-relaxed Sinkhorn algorithm}
%\label{SOR}
%\begin{algorithmic}
%\REQUIRE $\mu^1\in \IR^{n_1}$, $\mu^2\in \IR^{n_2}$, $c\in \IR^{n_1\times n_2}_+$
%\STATE Set $a=\mathbf{1}_{n_1}$, $b=\mathbf{1}_{n_2}$, $\gamma^0=e^{-c/\epsilon}$, $\omega\in[1;2[$ and $\eta>0$
%\WHILE {$||a\otimes \gamma^0b-  \mu_1||>\eta$}
%\STATE $\tilde a=\mu_1\oslash (\gamma^0 b)$,  $a_\omega=a^{1-\omega}\otimes \tilde a^\omega$
%\STATE $\tilde b=\mu_2\oslash (^t\gamma^0  a_\omega)$
%\IF{ $\langle \mu^1,\varphi_\omega(a\oslash\tilde a) \rangle +\langle \mu^2,\varphi_\omega(b\oslash\tilde b)\rangle>0$}
%\STATE  $a=a_\omega$, $b=b^{1-\omega}\otimes \tilde b^\omega$\COMMENT{Over-relaxed Sinkhorn iteration}
%\ELSE
%\STATE $a=\tilde a$, $b=\mu_2\oslash (^t\gamma^0 a)$  \COMMENT{Sinkhorn iteration}
%\ENDIF
%\ENDWHILE
% \RETURN $\gamma=\diag(a)\gamma^0\diag(b)$
%\end{algorithmic}
%\end{algorithm}
In order to summarize the whole algorithm, we first reformulate the functions $\theta_k(\gamma)$ defined in \eqref{eq:theta_k} as $\theta_k(\gamma)=\Theta(\frac{A_k \gamma}{\mu^k})$ where 
$ \Theta(u) = \min(\max(1,\Theta^*(u)-\delta),\theta_0)$ and  $\Theta^*(u)=\sup \left\{\omega \in [1;2]  \mid \varphi_\omega\left(\min u \right) \ge 0 \right\}$.

>>>>

\begin{proposition}
	We argue that a good choice for the functions $\theta_k$ may be constructed as follows. Pick a target parameter $\theta_0 \in [1;2)$, and a small $\delta>0$.
	Define the functions $\Theta^*$ and $\Theta$ as:
	\begin{align}
	\Theta^*(u) &= \sup \left\{\omega \in [1;2]  \mid \varphi_\omega\left(\min u\right) \ge 0 \right\} ,\\
	\label{eq:Theta}
	\Theta(u) &= \min(\max(1,\Theta^*(u)-\delta),\theta_0),
	\end{align}
	where $\min u$ denotes the smallest coordinate of the vector $u$.
	Then the function
	\begin{equation}
	\label{eq:theta_k}
	\theta_k(\gamma) =\Theta\left (\frac{A_k \gamma}{\mu^k}\right)
	\end{equation}
	is continuous, and verifies condition (\ref{eq:cond_theta_k}).
\end{proposition}
This construction has several advantages.
One is that it allows to choose arbitrarily the parameter $\theta_0$ that will be used asymptotically. A heuristic for this choice is proposed in section \ref{section:local}.
But it is also an easy approach to having an adaptive method, as the approximation of $\theta_k^*$ has a negligible cost (it only require to solve a one dimensional problem that depends on the smallest value of $\frac{A_k \gamma}{\mu^k}$, and the corresponding function to invert is decreasing and concave).
We will show later on that this is often an excellent choice in practice.
\begin{proof}
	Since the partial derivative of $\frac{\partial\varphi_\omega(x)}{\partial\omega}$ is nonzero for any $x<1$, the implicit function theorem proves the continuity of $\Theta^*$.
	The function $\Theta^*\left(\frac{A_k \gamma}{\mu^k}\right)$ is such that every term in relation (\ref{eq:kl_diff_scal}) is non-negative.
	Therefore, by lemma \ref{lemma:F_P_theta}, using this parameter minus $\delta$ ensures the strong decrease (\ref{eq:cond_theta_k}) of the Lyapunov function.
	It is no problem to constrain it to $[1,\theta_0]$, as using 1 always verifies the inequality (\ref{eq:cond_theta_k}), and using $\theta_0$ when it is smaller than $\Theta^*\left(\frac{A_k \gamma}{\mu^k}\right)$ does as well.
\end{proof}

@@

\renewcommand{\algorithmiccomment}[1]{\hfill\bgroup(#1)\egroup}
%\begin{algorithm}
%\caption{Over-relaxed Sinkhorn algorithm}
%\label{SOR}
%\begin{algorithmic}
%\REQUIRE $\mu^1\in \IR^{n_1}$, $\mu^2\in \IR^{n_2}$, $c\in \IR^{n_1\times n_2}_+$
%\STATE Set $a=\mathbf{1}_{n_1}$, $b=\mathbf{1}_{n_2}$, $\gamma^0=e^{-c/\epsilon}$, $\omega\in[1;2[$ and $\eta>0$
%\WHILE {$||a\otimes \gamma^0b-  \mu_1||>\eta$}
%\STATE $\tilde a=\mu_1\oslash (\gamma^0 b)$,  $a_\omega=a^{1-\omega}\otimes \tilde a^\omega$
%\STATE $\tilde b=\mu_2\oslash (^t\gamma^0  a_\omega)$
%\IF{ $\langle \mu^1,\varphi_\omega(a\oslash\tilde a) \rangle +\langle \mu^2,\varphi_\omega(b\oslash\tilde b)\rangle>0$}
%\STATE  $a=a_\omega$, $b=b^{1-\omega}\otimes \tilde b^\omega$\COMMENT{Over-relaxed Sinkhorn iteration}
%\ELSE
%\STATE $a=\tilde a$, $b=\mu_2\oslash (^t\gamma^0 a)$  \COMMENT{Sinkhorn iteration}
%\ENDIF
%\ENDWHILE
% \RETURN $\gamma=\diag(a)\gamma^0\diag(b)$
%\end{algorithmic}
%\end{algorithm}

This algorithm uses the function $\Theta$ defined in (\ref{eq:Theta}).
